---
params:
  data_dir: ~/Desktop/combination_compare/new_veneto/BRD-K62391742_BRD-K00005263
  comp: BRD-K62391742_BRD-K00005263
  qc_path: ~/Desktop/combination_compare/new_veneto/MTS016_VALIDATION_COMPOUND_INTERNAL_QC_TABLE.csv
output:
  html_document:
    toc: false
    theme: paper
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10)

# packages
library(tidyverse)
library(magrittr)
library(ggthemes)
library(plotly)
library(crosstalk)
library(htmltools)
library(DT)
library(heatmaply)
library(viridis)
library(splitstackshape)

# theme
theme_set(theme_bw())

# push NAs to bottom in tables
options(DT.TOJSON_ARGS = list(na = "string"))
```

```{r load data, include=FALSE}
# file paths
lfc_path <- list.files(params$data_dir, "LEVEL5_LFC_COMBAT", full.names = T)
drc_path <- list.files(params$data_dir, "DRC_TABLE", full.names = T)

# read in LFC
if (length(lfc_path) == 1) {
  LFC <- data.table::fread(lfc_path)
  combo_lfc <- LFC
} else {
  message("No LFC file found")
  quit(save = "no")
}

n_combos = length(LFC$pert_dose %>% unique())
synergy_height = ceiling(n_combos / 8) * 5

# read in DRC if available
if (length(drc_path) == 1) {
  DRC <- data.table::fread(drc_path) %>%
    dplyr::mutate(ec50 = as.numeric(ec50))
  combo_drc <- DRC
} else {
  DRC <- NULL
}

# read in QC table
plates <- LFC$pert_plate %>% unique()
qc_table <- data.table::fread(params$qc_path) %>%
  dplyr::filter(pert_plate %in% plates)

pert_iname <- toupper(dplyr::distinct(LFC, pert_iname)$pert_iname)

# biomarkers
corr_path <- list.files(params$data_dir, "continuous_associations", full.names = T)
if (length(corr_path) == 1) {
  correlations <- data.table::fread(corr_path) %>%
    dplyr::filter(q.val <= 0.1, feature_type %in% c("GE", "XPR", "REP")) %>%
    dplyr::mutate(feature = word(feature, 2, -1, sep = fixed("_")))
}
model_path <- list.files(params$data_dir, "model_table", full.names = T)
if (length(model_path) == 1) model_table <- data.table::fread(model_path)
rf_path <- list.files(params$data_dir, "RF_table", full.names = T)
if (length(rf_path) == 1) rf_table <- data.table::fread(rf_path)

# constituent compound info
constituents <- stringr::str_split(params$comp, pattern = fixed("_"))[[1]]
plate_dir <- dirname(params$data_dir)

# read in constitutent compound data
for (const in constituents) {
  const_path <- file.path(plate_dir, const)
  if (!dir.exists(const_path)) {
    warning("No data directory for anchor or test compounds. Unable to generate combination report.")
    quit(save = "no")
  }
  
  # DRC and LFC
  const_lfc_path <- list.files(const_path, "LEVEL5_LFC_COMBAT", full.names = T)
  const_drc_path <- list.files(const_path, "DRC_TABLE", full.names = T)
  if (length(const_lfc_path) == 1) {
    LFC %<>% dplyr::bind_rows(
      data.table::fread(const_lfc_path) %>%
        dplyr::mutate(pert_dose = as.character(pert_dose))
    )
  }
  if (length(const_drc_path) == 1) {
    DRC %<>% dplyr::bind_rows(
      data.table::fread(const_drc_path) %>%
        dplyr::mutate(ec50 = as.numeric(ec50))
    )
  }
  
  # correlations and random forest
  const_corr_path <- list.files(const_path, "continuous_associations", full.names = T)
  const_model_path <- list.files(const_path, "model_table", full.names = T)
  const_rf_path <- list.files(const_path, "RF_table", full.names = T)
  
  if (length(const_corr_path) == 1) {
    correlations %<>% dplyr::bind_rows(
      data.table::fread(const_corr_path) %>%
        dplyr::mutate(pert_dose = as.character(pert_dose)) %>%
        dplyr::mutate(feature = word(feature, 2, -1, sep = fixed("_")))
    )
  }
  if (length(const_model_path) == 1) {
    model_table %<>% dplyr::bind_rows(
      data.table::fread(const_model_path) %>%
        dplyr::mutate(pert_dose = as.character(pert_dose))
    )
  }
  if (length(const_rf_path) == 1) {
    rf_table %<>% dplyr::bind_rows(
      data.table::fread(const_rf_path) %>%
        dplyr::mutate(pert_dose = as.character(pert_dose)) %>%
        dplyr::mutate(feature = word(feature, 2, -1, sep = fixed("_")))
    )
  }
}
```

---
title:  "`r paste(pert_iname, 'combination report')`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Overview

This is the combination report for the combination `r pert_iname`. The doses at which this combination was run, as well as the doses of each single agent, are outlined in the table below

```{r overview}
# split to show components of compounds
summary_tab <- LFC %>%
  dplyr::distinct(pert_iname, pert_dose, pert_idose) %>%
  dplyr::rename(Compound = pert_iname, Dose = pert_dose, "Dose with units" = pert_idose) %>%
  splitstackshape::cSplit(splitCols = c("Compound", "Dose", "Dose with units"),
                          sep = "|", fixed = T,
                          direction = "wide", drop = T) %>%
  dplyr::arrange(Compound_2, Compound_1, Dose_1)

colnames(summary_tab) <- stringr::str_replace_all(colnames(summary_tab), "_", " ")

DT::datatable(summary_tab, style = "bootstrap",
              options = list(dom = "t", ordering = TRUE,
                             lengthChange = FALSE, scrollY = "300px", paging = FALSE))
```

# Sensitivity comparison {.tabset .tabset-pills}

## Overview

This section compares the activity of combination with constituents based on log-fold change and AUC.

The first plot shows synergy score (x-axis) and the significance of that score (y-axis). Synergy score is calculated as $LFC_a + LFC_b - LFC_c$ where $LFC_a$ and $LFC_b$ are the log-fold change in each constituent compound and $LFC_c$ is the log-fold change in the combination. Large positive values indicate synergy while large negative values indicate antagonism. Significance metrics, $q$ values, are obtained by normalizing the synergy score to the variance of the cell line abundance in DMSO, calculating a $p$ value and correcting it for multiple hypothesis testing.

The second plot shows the AUC of the combination versus the Bliss expected AUC. Each value is calculated by taking the mean fold-change across doses with a given anchor dose as a proxy for AUC. For example:
$$AUC_{bliss} \approx \frac{1}{n}\sum_{i=1}^nFC_{a_i}FC_b$$
$$AUC_{combo} \approx \frac{1}{n}\sum_{i=1}^nFC_{c_i}$$
The same is repeated for compound $b$ if there are more than 4 doses.

## Synergy by dose

```{r synergy scoring, fig.height=synergy_height}
# non-combination values
solo_lfc <- dplyr::anti_join(LFC, combo_lfc)

# split combination perts
spread_lfc <- combo_lfc %>%
  dplyr::inner_join(qc_table %>%
                      dplyr::group_by(ccle_name, culture) %>%
                      dplyr::summarise(ctl_vehicle_mad = median(ctl_vehicle_mad),
                                       .groups = "drop")) %>%  # QC info
  dplyr::distinct(pert_iname, pert_dose, ccle_name, culture, LFC_cb, ctl_vehicle_mad) %>%
  splitstackshape::cSplit(splitCols = c("pert_iname", "pert_dose"),
                          sep = "|", fixed = T,
                          direction = "wide", drop = F, type.convert = T)

# add columns for constituents to spread table
for (i in 1:length(constituents)) {
  comp <- constituents[[i]]
  df <- solo_lfc %>%
    dplyr::filter(toupper(pert_id) == toupper(comp)) %>%
    dplyr::mutate(pert_dose = as.numeric(pert_dose)) %>%
    dplyr::distinct(pert_iname, pert_dose, ccle_name, culture, LFC_cb)
  
  new_cols <- paste(colnames(df), i, sep = "_")
  colnames(df) <- c(colnames(df)[1:2], new_cols[3:length(new_cols)])
  
  spread_lfc %<>% dplyr::left_join(df)
}

# bliss expected fold change (capped FCs at 1)
B <- apply(2^spread_lfc[,11:ncol(spread_lfc)], 1, function(x) prod(pmin(x, 1)))
spread_lfc$S <- rowSums(spread_lfc[,11:ncol(spread_lfc)]) - spread_lfc$LFC_cb
spread_lfc %<>%
  dplyr::mutate(Z  = S / ctl_vehicle_mad / sqrt(pi / 2 )) %>%
  dplyr::mutate(p = 2*pnorm(abs(Z), lower.tail = F)) %>%
  dplyr::mutate(q = p.adjust(p, method = "BH")) %>%
  dplyr::arrange(pert_dose_1)

if (all(is.na(spread_lfc$S))) {
  warning("No matching data from combination constituents alone, unable to generate combination report.")
  quit(save = "no")
}

max_s <- ceiling(max(spread_lfc$S))
min_s <- floor(min(spread_lfc$S))
max_q <- ceiling(max(-log10(spread_lfc$q)))

combos <- spread_lfc %>% dplyr::distinct(pert_dose)

plots <- list()
for (i in 1:nrow(combos)) {
  synergy_plot <- highlight_key(spread_lfc %>% dplyr::inner_join(combos[i,]),
                                ~ccle_name, group = "Synergy scores")
  p <- plot_ly(synergy_plot) %>%
    add_trace(x = ~S, y = ~-log10(q), type = "scatter", mode = "markers",
              hoverinfo = "text", text = ~ccle_name, opacity = 0.5) %>%
    layout(xaxis = list(range = c(min_s, max_s)),
           yaxis = list(range = c(0, max_q)),
           annotations = list(x = 0.5, y = 1,
                              text = combos$pert_dose[[i]],
                              align = "center", showarrow = F, xref="paper", yref="paper",
                              xanchor = "center", yanchor = "top"),
           hovermode = "closest",
           showlegend = F) %>%
    highlight(on = "plotly_click", off = "plotly_doubleclick",
              dynamic = FALSE, selectize = FALSE, persistent = FALSE,
              defaultValues = "", selected = attrs_selected(opacity = 1))
  plots[[i]] <- p
}
subplot(plots, nrows = max(1, ceiling(length(plots)/4)), shareX = F, shareY = F) %>%
  layout(title = pert_iname,
         annotations = list(
           list(x = 0, xshift = -50, y = 0.5,
                textangle = 270, text = "-log10(q)",
                font = list(size = 14),
                showarrow = F, xref='paper', yref='paper',
                size = 48),
           list(x = 0.5, y = 0, yshift = -40,
                text = "Synergy score",
                font = list(size = 14),
                showarrow = F, xref='paper', yref='paper',
                size = 48)
         ))
```

## AUC plots {.tabset .tabset-fade}

```{r auc, include=F, fig.height=10}
# calculate fold-change of combination
spread_lfc$FC <- pmin(2^spread_lfc$LFC_cb, 1)
spread_lfc$B <- B

# group by dose 2 to calculate AUC and expected for comp 1
pseudo_auc_1 <- spread_lfc %>%
  dplyr::select(ccle_name, culture,
                pert_iname_1, pert_dose_1, pert_dose_2, FC, B) %>%
  dplyr::rename(pert_iname = pert_iname_1) %>%
  dplyr::group_by(ccle_name, culture, pert_iname, pert_dose_2) %>%
  dplyr::mutate(n = n()) %>%
  dplyr::summarise_all(mean) %>%
  dplyr::ungroup() %>%
  dplyr::select(-pert_dose_1) %>%
  dplyr::rename(pert_dose = pert_dose_2) %>%
  dplyr::filter(n >= 4)

# group by dose 1 to calculate AUC and expected for comp 2
pseudo_auc_2 <- spread_lfc %>%
  dplyr::select(ccle_name, culture,
                pert_iname_2, pert_dose_1, pert_dose_2,
                FC, B) %>%
  dplyr::rename(pert_iname = pert_iname_2) %>%
  dplyr::group_by(ccle_name, culture, pert_iname, pert_dose_1) %>%
  dplyr::mutate(n = n()) %>%
  dplyr::summarise_all(mean) %>%
  dplyr::ungroup() %>%
  dplyr::select(-pert_dose_2) %>%
  dplyr::rename(pert_dose = pert_dose_1) %>%
  dplyr::filter(n >= 4)

pseudo_auc <- dplyr::bind_rows(pseudo_auc_1, pseudo_auc_2) %>%
  dplyr::mutate(cell_id = paste(ccle_name, culture))
comps <- pseudo_auc$pert_iname %>% unique()

super_plots <- list()
for (i in 1:length(comps)) {
  comp = comps[[i]]
  df <- pseudo_auc %>% dplyr::filter(pert_iname == comp)
  doses <- df$pert_dose %>% unique()
  
  plots = list()
  for (j in 1:length(doses)) {
    df <- highlight_key(pseudo_auc %>% dplyr::filter(pert_dose == doses[[j]]),
                        ~ccle_name, group = comp)
    
    p <- plot_ly() %>%
      add_trace(data = df,
                x = ~B, y = ~FC,
                type = "scatter", mode = "markers", opacity = 0.5,
                hoverinfo = "text", text = ~cell_id) %>%
      layout(xaxis = list(title = "Bliss expected AUC", range = c(0, 1)),
             yaxis = list(title = "AUC", range = c(0, 1)),
             hovermode = "closest",
             shapes = list(type = "line", y0=0, y1= 1, x0=0, x1= 1,
                           line = list(dash = "dash", layer = "below")),
             annotations = list(x = 0.5, y = 1,
                              text = doses[[j]],
                              align = "center", showarrow = F, xref="paper", yref="paper",
                              xanchor = "center", yanchor = "top"),
             showlegend = F) %>%
      highlight(on = "plotly_selected", off = "plotly_deselect",
                dynamic = FALSE, selectize = TRUE, persistent = FALSE,
                opacityDim = 1, defaultValues = "",
                selected = attrs_selected(opacity = 1))
    
    plots[[j]] <- p
  }
  
  super_plots[[i]] <- subplot(plots, nrows = max(1, ceiling(length(plots)/2)), shareX = T, shareY = T)
}
```

```{r auc plots, fig.height=10, results='asis'}
for (i in 1:length(super_plots)) {
  cat(sprintf('\n\n### %s \n\n', comps[[i]], '\n\n'))
  cat(htmltools::knit_print.shiny.tag(super_plots[[i]]))
  cat("\n\n")
}
```


# Continuous associations {.tabset .tabset-pills}

## Overview

Results from correlation with various DepMap data sets. For simplicity only correlations with AUC or a single anchor dose are visualized. To highlight a feature either click on a point or select a row from the table below the plot. Selecting a feature will highlight it in every plot in which it appears. Only features with a $q$ value less than $0.1$ and in the top 500 features will be visualized. If no plot appears for a particular constituent or the combination then there were no features meeting threshold for that case.

```{r biomarker pre, include=F}
single_dose <- LFC %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n_doses = n_distinct(pert_dose), .groups = "drop") %>%
  dplyr::filter(n_doses == 1) %>% .$pert_iname

sub_correlations <- correlations %>%
  dplyr::filter(pert_dose == "log2.auc" | pert_iname %in% single_dose,
                is.finite(coef), is.finite(q.val),
                q.val < 0.1, rank <= 250) %>%
  dplyr::mutate(added_doses = ifelse(is.na(added_doses), "None", added_doses),
                added_compounds = ifelse(is.na(added_compounds), "None", added_compounds))
sub_models <- model_table %>%
  dplyr::filter(pert_dose == "log2.auc" | pert_iname %in% single_dose) %>%
  dplyr::mutate(added_doses = ifelse(is.na(added_doses), "None", added_doses),
                added_compounds = ifelse(is.na(added_compounds), "None", added_compounds))
sub_rf <- rf_table %>%
  dplyr::filter(pert_dose == "log2.auc" | pert_iname %in% single_dose) %>%
  dplyr::filter(rank <= 100) %>%
  dplyr::mutate(added_doses = ifelse(is.na(added_doses), "None", added_doses),
                added_compounds = ifelse(is.na(added_compounds), "None", added_compounds))
```

## Gene expression {.tabset .tabset-fade}

### Plots

```{r gene expression}
features_ge <- sub_correlations %>%
  dplyr::filter(feature_type == "GE")

combos <- features_ge %>% dplyr::distinct(pert_iname, added_compounds)

plots <- list()
if (nrow(combos) < 1) {
  print("No significant correlations to plot")
} else {
  for (i in 1:nrow(combos)) {
    volcano_combo <- highlight_key(features_ge %>% dplyr::inner_join(combos[i,]),
                                   ~feature, group = "GE correlations")
    p <- plot_ly(volcano_combo) %>%
      add_trace(x = ~coef, y = ~-log10(q.val), type = "scatter", mode = "markers",
                hoverinfo = "text", text = ~feature, opacity = 0.3, color = ~added_doses,
                showlegend = T) %>%
      layout(xaxis = list(title = "Correlation coefficient"),
             yaxis = list(title = "-log10(q-value)"), hovermode = "closest",
             annotations = list(x = 0.5, y = 1,
                                text = paste(combos[i, "pert_iname"], "+", combos[i, "added_compounds"]),
                                align = "center", showarrow = F, xref="paper", yref="paper",
                                xanchor = "center", yanchor = "top")) %>%
      highlight(on = "plotly_click", off = "plotly_doubleclick",
                dynamic = FALSE, selectize = FALSE, persistent = FALSE,
                opacityDim = 1, defaultValues = "",
                selected = attrs_selected(opacity = 1))
    plots[[i]] <- p
  }
  subplot(plots, nrows = max(1, ceiling(length(plots)/2)), shareX = T, shareY = T)
}
features_combo <- features_ge %>%
  dplyr::select(pert_iname, pert_dose, added_compounds, added_doses, feature, coef, q.val) %>%
  dplyr::mutate(coef = signif(coef, digits = 3), q.val = signif(q.val, digits = 3))
features_combo_dt <- highlight_key(features_combo, ~feature, group = "GE correlations")
dt <- DT::datatable(features_combo_dt, style="bootstrap", width="100%",
                    options = list(dom = "t", lengthChange = FALSE, scrollY = "300px",
                                   scrollX = "300px", paging = FALSE),
                    filter = "top",
                    colnames = c("Test Compound", "Dose", "Anchor Compound(s)", "Anchor Dose(s)", "Feature", "Correlation Coefficient", "q-value"))
dt
```

### Heatmap

Rank of gene expression features that appear in the top 10 of one of the conditions. `NA` value indicates a rank below 100 for that feature in that condition.

```{r expression heatmap, fig.height=5}
top_feats <- features_ge %>% dplyr::filter(rank <= 10) %>% .$feature %>% unique()
top_tab <- features_ge %>%
  dplyr::filter(feature %in% top_feats) %>%
  reshape2::acast(pert_iname + added_compounds + added_doses ~ feature, value.var = "rank", fun.aggregate = min)
top_tab[top_tab > 100] <- NA
top_tab <- top_tab[, order(apply(top_tab, 2, function(x) mean(x, na.rm = T)))]
if (is.null(dim(top_tab))) {
  print("Only one condition with relevant feaatures, no heatmap to display")
} else {
  heatmaply::heatmaply(top_tab, Rowv = F, Colv = F, colors = viridis(n = 256, direction = -1),
                       key.title = "Rank", label_names = c("Compound", "Gene", "Rank"))
}
```

## CRISPR dependency {.tabset .tabset-fade}

### Plots

To highlight a gene either click on a point or select a row from the table below. Selecting a gene will highlight it in every plot in which it appears.

```{r crispr}
features_xpr <- sub_correlations %>%
  dplyr::filter(feature_type == "XPR")

combos <- features_xpr %>% dplyr::distinct(pert_iname, added_compounds)

plots <- list()
if (nrow(combos) < 1) {
  print("No significant correlations to plot")
} else {
  for (i in 1:nrow(combos)) {
    volcano_combo <- highlight_key(features_xpr %>% dplyr::inner_join(combos[i,]),
                                   ~feature, group = "XPR correlations")
    p <- plot_ly(volcano_combo) %>%
      add_trace(x = ~coef, y = ~-log10(q.val), type = "scatter", mode = "markers",
                hoverinfo = "text", text = ~feature, opacity = 0.3, color = ~added_doses,
                showlegend = T) %>%
      layout(xaxis = list(title = "Correlation coefficient"),
             yaxis = list(title = "-log10(q-value)"), hovermode = "closest",
             annotations = list(x = 0.5, y = 1,
                                text = paste(combos[i, "pert_iname"], "+", combos[i, "added_compounds"]),
                                align = "center", showarrow = F, xref="paper", yref="paper",
                                xanchor = "center", yanchor = "top")) %>%
      highlight(on = "plotly_click", off = "plotly_doubleclick",
                dynamic = FALSE, selectize = FALSE, persistent = FALSE,
                opacityDim = 1, defaultValues = "",
                selected = attrs_selected(opacity = 1))
    plots[[i]] <- p
  }
  subplot(plots, nrows = max(1, ceiling(length(plots)/2)), shareX = T, shareY = T)
}
features_combo <- features_xpr %>%
  dplyr::select(pert_iname, pert_dose, added_compounds, added_doses, feature, coef, q.val) %>%
  dplyr::mutate(coef = signif(coef, digits = 3), q.val = signif(q.val, digits = 3))
features_combo_dt <- highlight_key(features_combo, ~feature, group = "XPR correlations")
dt <- DT::datatable(features_combo_dt, style="bootstrap", width="100%",
                    options = list(dom = "t", lengthChange = FALSE, scrollY = "300px",
                                   scrollX = "300px", paging = FALSE),
                    filter = "top",
                    colnames = c("Test Compound", "Dose", "Anchor Compound(s)", "Anchor Dose(s)", "Feature", "Correlation Coefficient", "q-value"))
dt
```

### Heatmap

Rank of CRISPR dependency features that appear in the top 10 of one of the conditions. `NA` value indicates a rank below 100 for that feature in that condition.

```{r crispr heatmap, fig.height=5}
top_feats <- features_xpr %>% dplyr::filter(rank <= 10) %>% .$feature %>% unique()
top_tab <- features_xpr %>%
  dplyr::filter(feature %in% top_feats) %>%
  reshape2::acast(pert_iname + added_compounds + added_doses ~ feature, value.var = "rank", fun.aggregate = min)
top_tab[top_tab > 100] <- NA
top_tab <- top_tab[, order(apply(top_tab, 2, function(x) mean(x, na.rm = T)))]
if (is.null(dim(top_tab))) {
  print("Only one condition with relevant feaatures, no heatmap to display")
} else {
  heatmaply::heatmaply(top_tab, Rowv = F, Colv = F, colors = viridis(n = 256, direction = -1),
                       key.title = "Rank", label_names = c("Compound", "Gene", "Rank"))
}
```

## Repurposing compounds {.tabset .tabset-fade}

### Plots

To highlight a compound either click on a point or select a row from the table below. Selecting a compound will highlight it in every plot in which it appears.

```{r rep}
features_rep <- sub_correlations %>%
  dplyr::filter(feature_type == "REP")

combos <- features_rep %>% dplyr::distinct(pert_iname, added_compounds)
plots <- list()
if (nrow(combos) < 1) {
  print("No significant correlations to plot")
} else {
  for (i in 1:nrow(combos)) {
    volcano_combo <- highlight_key(features_rep %>% dplyr::inner_join(combos[i,]),
                                   ~feature, group = "REP correlations")
    p <- plot_ly(volcano_combo) %>%
      add_trace(x = ~coef, y = ~-log10(q.val), type = "scatter", mode = "markers",
                hoverinfo = "text", text = ~feature, opacity = 0.3, color = ~added_doses,
                showlegend = T) %>%
      layout(xaxis = list(title = "Correlation coefficient"),
             yaxis = list(title = "-log10(q-value)"), hovermode = "closest",
             annotations = list(x = 0.5, y = 1,
                                text = paste(combos[i, "pert_iname"], "+", combos[i, "added_compounds"]),
                                align = "center", showarrow = F, xref="paper", yref="paper",
                                xanchor = "center", yanchor = "top")) %>%
      highlight(on = "plotly_click", off = "plotly_doubleclick",
                dynamic = FALSE, selectize = FALSE, persistent = FALSE,
                opacityDim = 1, defaultValues = "",
                selected = attrs_selected(opacity = 1))
    plots[[i]] <- p
  }
  subplot(plots, nrows = max(1, ceiling(length(plots)/2)), shareX = T, shareY = T)
}
features_combo <- features_rep %>%
  dplyr::select(pert_iname, pert_dose, added_compounds, added_doses, feature, coef, q.val) %>%
  dplyr::mutate(coef = signif(coef, digits = 3), q.val = signif(q.val, digits = 3))
features_combo_dt <- highlight_key(features_combo, ~feature, group = "REP correlations")
dt <- DT::datatable(features_combo_dt, style="bootstrap", width="100%",
                    options = list(dom = "t", lengthChange = FALSE, scrollY = "300px",
                                   scrollX = "300px", paging = FALSE),
                    filter = "top",
                    colnames = c("Test Compound", "Dose", "Anchor Compound(s)", "Anchor Dose(s)", "Feature", "Correlation Coefficient", "q-value"))
dt
```

### Heatmap

Rank of repurposing features that appear in the top 10 of one of the conditions. `NA` value indicates a rank below 100 for that feature in that condition.

```{r rep heatmap, fig.height=5}
top_feats <- features_rep %>% dplyr::filter(rank <= 10) %>% .$feature %>% unique()
top_tab <- features_rep %>%
  dplyr::filter(feature %in% top_feats) %>%
  reshape2::acast(pert_iname + added_compounds + added_doses ~ feature, value.var = "rank", fun.aggregate = min)
top_tab[top_tab > 100] <- NA
top_tab <- top_tab[, order(apply(top_tab, 2, function(x) mean(x, na.rm = T)))]
if (is.null(dim(top_tab))) {
  print("Only one condition with relevant feaatures, no heatmap to display")
} else {
  heatmaply::heatmaply(top_tab, Rowv = F, Colv = F, colors = viridis(n = 256, direction = -1),
                       key.title = "Rank", label_names = c("Compound", "Gene", "Rank"))
}
```

# Random forest models {.tabset .tabset-pills}

## Overview

Results from random forests fit to predict response based on multi-omics datasets. For simplicity only models predicting AUC or a single anchor dose are visualized.. For more information on how random forests are fit see a standard PRISM report.

## Results

These plots show the accuracy of predictive models (left) and the top 100 features used ranked by importance in each model (right)

``` {r multi ccle, fig.height=8}
p1 <- sub_models %>%
  dplyr::mutate(condition = paste(pert_iname, added_compounds, added_doses)) %>%
  plot_ly() %>%
  add_trace(x = ~model, y = ~R2, color = ~condition,
            type = "bar", hoverinfo = "text", text = ~R2) %>%
  layout(showlegend = FALSE, xaxis = list(title = "Model"))
p2 <- sub_rf %>%
  dplyr::mutate(condition = paste(pert_iname, added_compounds, added_doses)) %>%
  plot_ly() %>%
  add_trace(x = ~rank, y = ~RF.imp.mean, color = ~condition, symbol = ~model,
            type = "scatter", mode = "markers",
            hoverinfo = "text", text = ~feature) %>%
  layout(xaxis = list(type = "log", title = "Rank"),
         yaxis = list(title = "Feature importance"))

df <- sub_models %>% dplyr::inner_join(sub_rf) %>%
  dplyr::select(pert_iname, pert_dose, added_compounds, added_doses, model, R2, PearsonScore,
                feature, RF.imp.mean) %>%
  dplyr::mutate_at(vars(R2, PearsonScore, RF.imp.mean),
                   funs(signif(., 3)))

dt <- DT::datatable(df, style="bootstrap", width="100%",
                    options = list(dom = "t", lengthChange = FALSE, scrollY = "300px",
                                   scrollX = "300px", paging = FALSE,
                                   columnDefs = list(list(className = 'dt-center', targets = "_all"))),
                    filter = "top",
                    colnames = c("Test Compound", "Dose", "Anchor Compound(s)", "Anchor Dose(s)", "Model", "R2", "Pearson Score", "Feature", "Feature Importance"),)

subplot(style(p1, showlegend = FALSE), p2, widths = c(0.2, 0.8),
        titleX = T, titleY = T, margin = 0.05)
dt
```
