---
params:
  data_dir: ~/Desktop/MTS016
  screen_name: MTS016
title: "`r params$screen_name` QC Report"
author: "Andrew Boghossian"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: false
    theme: paper
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width = 10)

# load libraries
library(magrittr)
library(reshape2)
library(data.table)
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(mousetrap)
library(uwot)
library(DT)
library(plotly)
library(ggExtra)
library(heatmaply)

# load QC data (filter to 120H data (exclude base etc.))
qc_path <- list.files(params$data_dir, "QC_TABLE", full.names = T)
qc_table <- data.table::fread(qc_path) %>%
  dplyr::filter(pert_time == "120H")

# load level 4 data
level4_path <- list.files(params$data_dir, "LEVEL4_LFC_COMBAT", full.names = T)
if (length(level4_path) == 1) {
  level4_table <- data.table::fread(level4_path) %>%
    dplyr::filter(pert_time == "120H")
} else {
  level4_path <- list.files(params$data_dir, "LEVEL4_LFC", full.names = T)
  level4_table <- data.table::fread(level4_path) %>%
    dplyr::filter(pert_time == "120H") %>%
    dplyr::rename(LFC_cb = LFC)
}

# load LFC collapsed data (level 5)
lfc_path <- list.files(params$data_dir, "LEVEL5_LFC_COMBAT", full.names = T)
if (length(lfc_path) == 1) {
  lfc_table <- data.table::fread(lfc_path) %>%
    dplyr::filter(pert_time == "120H")
} else {
  lfc_path <- list.files(params$data_dir, "LEVEL5_LFC", full.names = T)
  lfc_table <- data.table::fread(lfc_path) %>%
    dplyr::filter(pert_time != "0H") %>%
    dplyr::rename(LFC_cb = LFC)
}

# push NAs to bottom in tables
options(DT.TOJSON_ARGS = list(na = "string"))

# plot theme
theme_set(theme_light())

mycolors <- tableau_color_pal()(10)
mycolors <- colorRampPalette(mycolors)(length(qc_table$prism_replicate %>% unique()))
```

## Overall performance {.tabset .tabset-fade}

### Included lines by plate

Number of cell lines included in LFC and downstream analysis per plate (must pass >50% of replicates to be included).

```{r summary}
# sum number lines passing >= 2 replicates on each plate/culture (bar chart)
qc_table %>%
  dplyr::group_by(ccle_name, culture, pool_id, pert_plate) %>%
  dplyr::summarise(pass = sum(pass) > n()/2, .groups = "drop") %>%
  dplyr::group_by(pass, culture, pert_plate) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  dplyr::group_by(pert_plate, culture) %>%
  dplyr::mutate(total = sum(count)) %>%
  dplyr::ungroup() %>%
  ggplot(aes(x = pass, y = count/total * 100)) +
  geom_bar(aes(fill = pass), stat = "identity") +
  facet_grid(culture ~ pert_plate) +
  geom_text(aes(label = count)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_fivethirtyeight(limits = c("TRUE", "FALSE")) +
  labs(fill = "Pass", y = "Percent")
```

### Passing lines by replicate

Number of cell lines passing QC per replicate. Must have dynamic range > $\log_2(0.3) \approx 1.74$ and error rate $\leq$ 0.05.

```{r rep summary, fig.height=10}
# same as above on replicate level
qc_table %>%
  dplyr::group_by(pass, culture, pert_plate, prism_replicate) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  dplyr::group_by(pert_plate, prism_replicate, culture) %>%
  dplyr::mutate(total = sum(count)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(replicate = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_")))) %>%
  ggplot(aes(x = pass, y = count/total * 100)) +
  geom_bar(aes(fill = pass), stat = "identity") +
  facet_wrap(culture ~ replicate) +
  geom_text(aes(label = count)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_fivethirtyeight(limits = c("TRUE", "FALSE")) +
  labs(fill = "Pass", y = "Percent")
```

### Passing lines by pool

Proportion of cell lines passing QC per pool per replicate. Must have dynamic range > $\log_2(0.3) \approx 1.74$ and error rate $\leq$ 0.05.

```{r pool summary, fig.height=8}
# same as above on pool x replicate level
pool_matrix <- qc_table %>%
  dplyr::group_by(pert_plate, prism_replicate, pool_id, culture) %>%
  dplyr::summarise(count = n(),
                   passing = sum(pass),
                   .groups = "drop") %>%
  dplyr::mutate(passing_prop = passing/count) %>%
  dplyr::mutate(replicate = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_"))),
                replicate = word(replicate, 1, sep = fixed("."))) %>%
  reshape2::acast(pool_id + culture ~ replicate, value.var = "passing_prop")

heatmaply::heatmaply(pool_matrix, Rowv = F, Colv = F, limits = c(0, 1))
```

### Removed lines

Table of cell lines that were removed from each plate (must fail on $\geq$ 50% of replicates). Sorted by number of plates removed from.

```{r failures}
# find lines that are failing on every replicate of a plate
# show in table (and how many plates they are failing on)
dt <- qc_table %>%
  dplyr::group_by(ccle_name, culture, pool_id, pert_plate) %>%
  dplyr::summarise(pass = max(pass), .groups = "drop") %>%
  dplyr::filter(pass == 0) %>%
  dplyr::select(-pass) %>%
  dplyr::group_by(ccle_name, pool_id, culture) %>%
  dplyr::summarise(n_failed = n(),
                   failed_plates = paste(pert_plate, collapse = "|"),
                   .groups = "drop") %>%
  dplyr::arrange(desc(n_failed))

DT::datatable(dt, rownames = F, options = list(dom = "tp"), filter = "top",
              colnames = c("Cell Line", "Pool", "Culture", "Num. Failed plates", "Failed plates"))
```

## Dynamic range and error rate {.tabset .tabset-fade}

### Overview

Dynamic range versus error rate for each cell line on each replicate (points are cell lines). Points in red failed QC. Lines indicate passing thresholds for each metric.

```{r standard plot, fig.height=10}
# dynamic range versus error rate by replicate
qc_table %>%
  dplyr::mutate(replicate = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_")))) %>%
  ggplot(aes(x = dr, y = error_rate, color = pass)) +
  geom_vline(xintercept = -log2(0.3), linetype = "dashed") +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_point() +
  facet_wrap(culture ~ replicate) +
  scale_color_fivethirtyeight(limits = c("TRUE", "FALSE")) +
  labs(x = "Dynamic range", y = "Error rate", color = "Pass")
```

### Dynamic ranges

Cumulative distribution plot of cell line dynamic ranges on each replicate.

```{r dr, fig.height=8}
# cumulative distribution plot of dynamic range
qc_table %>%
  ggplot(aes(x = dr, color = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_"))))) +
  geom_vline(xintercept = -log2(0.3), linetype = "dashed") +
  stat_ecdf() +
  facet_wrap(culture ~ ., ncol = 1) +
  labs(color = "Replicate", x = "Dynamic range", y = "Proportion") +
  scale_color_manual(values = mycolors)
```

## Medians and MADs {.tabset .tabset-fade}

### DMSO

Median versus MAD logMFI in DMSO. Each point is a cell line. Points in red failed QC.

```{r medmad DMSO, fig.height=8}
# median vs MAD logMFI in DMSO
qc_table %>%
  dplyr::mutate(replicate = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_")))) %>%
  ggplot(aes(x = ctl_vehicle_md, y = ctl_vehicle_mad, color = pass)) +
  geom_point(alpha = 0.6) +
  facet_wrap(culture ~ replicate) +
  labs(x = "DMSO median", y = "DMSO MAD", color = "Pass") +
  scale_color_fivethirtyeight(limits = c("TRUE", "FALSE"))
```

### Bortezomib

Median versus MAD logMFI in bortezomib. Each point is a cell line. Points in red failed QC.

```{r medmad bort, fig.height=8}
# same as above in bortezomib
qc_table %>%
  dplyr::mutate(replicate = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_")))) %>%
  ggplot(aes(x = trt_poscon_md, y = trt_poscon_mad, color = pass)) +
  geom_point(alpha = 0.6) +
  facet_wrap(culture ~ replicate) +
  labs(x = "Bortezomib median", y = "Bortezomib MAD", color = "Pass") +
  scale_color_fivethirtyeight(limits = c("TRUE", "FALSE"))
```

### CDF plot

Cumulative distribution plot of cell line median logMFI in DMSO (solid lines) and bortezomib (dashed lines) on each replicate.

```{r medmad cdf, fig.height=8}
# cumulative distribution plot of dynamic range
qc_table %>%
  ggplot(aes(color = paste(pert_plate, word(prism_replicate, 4, sep = fixed("_"))))) +
  stat_ecdf(aes(x = ctl_vehicle_md)) +
  stat_ecdf(aes(x = trt_poscon_md), linetype = "dashed") +
  facet_wrap(culture ~ ., ncol = 1) +
  labs(color = "Replicate", x = "Normalized logMFI", y = "Proportion") +
  scale_color_manual(values = mycolors)
```

## Replicate correlations {.tabset .tabset-fade}

Correlation between replicates versus median correlation between replicates of perturbations across each replicate for each plate. Each point is a perturbation.

```{r rep cor, results='asis', fig.height=8}
plates <- level4_table$pert_plate %>% unique() %>% sort()
plots <- list()

for (i in 1:length(plates)) {
  plate <- plates[[i]]
  rep_cor_tab <- level4_table %>%
    dplyr::filter(pert_plate == plate, pert_type == "trt_cp", pool_id != "CTLBC")

  replicates <- rep_cor_tab$x_replicate %>% unique()
  if (length(replicates) < 2) {
    next
  }
  replicates_data <- list()
  for (j in 1:length(replicates)) {
    rep_val <- replicates[[j]]

    replicate_data <- rep_cor_tab %>%
      dplyr::filter(x_replicate == rep_val) %>%
      reshape2::acast(ccle_name + culture ~ pert_id + pert_dose + x_project_id,
                      value.var = "LFC_cb", fun.aggregate = mean)
    replicates_data[[j]] <- replicate_data
  }

  combos <- combn(1:length(replicates), m = 2)
  replicates_cor <- list()
  shared_perts <- Reduce(intersect, lapply(replicates_data, colnames))
  for (j in 1:ncol(combos)) {
    df1 <- replicates_data[[combos[1, j]]]
    df2 <- replicates_data[[combos[2, j]]]
    overlap <- dplyr::intersect(rownames(df1), rownames(df2))
    replicates_cor[[j]] <- cor(df1[overlap, shared_perts], df2[overlap, shared_perts]) %>% diag()
    names(replicates_cor)[[j]] <- paste("X", combos[, j], collapse = "", sep = "")
  }

  replicates_cor %<>% dplyr::bind_cols()
  replicates_cor$R0 <- apply(replicates_cor, 1, median)
  replicates_cor %<>%
    tidyr::gather("Rep", "cor", 1:3)

  p <- replicates_cor %>%
    ggplot(aes(x = R0, y = cor, color = Rep)) +
    geom_point() +
    geom_density_2d() +
    geom_abline() +
    scale_color_tableau() +
    scale_fill_tableau() +
    labs(x = "Median correlation", y = "Correlation", color = "Replicate correlation") +
    theme(legend.position = "bottom")

  p <- ggExtra::ggMarginal(p, type = "histogram", groupFill = T,
                           position = "identity")

  plots[[i]] <- p

  cat("\n\n### ")
  cat(plate)
  cat("\n\n")
  print(p, newpage = T)
  cat("\n\n")
}
```


## Cell line sensitivity

This plot shows the proportion of perturbations killing a given cell line on a given plate to 30% viability and the 90th percentile of LFC for that cell line. Each point is a cell line. Cell lines in the table are killed in 0, more that 40% of perturbations on a plate, or whose 90th percentile of LFC is > -1 (corresponding to viability of 50%).

```{r lfc pre, include=F}
# calculate proportion of perturbations killing to 30% viability
cl_summary <- lfc_table %>%
  dplyr::group_by(ccle_name, culture, pool_id, pert_plate) %>%
  dplyr::summarise(prop_killed = sum(LFC_cb < log2(0.3)) / n(),
                   ninety_perc = quantile(LFC_cb, probs = c(0.1)))
```

```{r cls, fig.height=8}
mycolors <- tableau_color_pal()(10)
mycolors <- colorRampPalette(mycolors)(length(cl_summary$pool_id %>% unique()))

cl_summary %>%
  ggplot(aes(x = ninety_perc, y = prop_killed, color = pool_id)) +
  geom_point(alpha = 0.6) +
  facet_wrap(pert_plate ~ culture) +
  labs(x = "90th percentile LFC", y = "Perturbations killing to 30%", color = "Pool") +
  scale_color_manual(values = mycolors)

DT::datatable(cl_summary %>%
                dplyr::filter(prop_killed == 0 | prop_killed > 0.4 | ninety_perc > -1) %>%
                dplyr::mutate(prop_killed = signif(prop_killed, 3),
                              ninety_perc = signif(ninety_perc, 3)),
              filter = "top", rownames = F, options = list(dom = "tp"),
              colnames = c("Cell line", "Culture", "Pool", "Plate", "Kiling proportion", "90th percentile LFC"))
```

## LFC comparisons {.tabset .tabset-fade}

Comparing log-fold changes of compounds based on the dose with the highest bimodality coefficient.

```{r comp compare, include = F}
# find dose for each pert/plate/project with highest bimodality coefficient
pert_bcs <- lfc_table %>%
  dplyr::group_by(pert_iname, pert_idose, x_project_id, pert_plate) %>%
  dplyr::summarise(bc = mousetrap::bimodality_coefficient(LFC_cb),
                   .groups = "drop") %>%
  dplyr::group_by(pert_iname, x_project_id, pert_plate) %>%
  dplyr::filter(bc == max(bc)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(pert = paste(pert_iname, pert_idose, x_project_id, pert_plate, sep = fixed("::")))

# correlate LFCs of each compound (single dose chosen above)
pert_cors <- lfc_table %>%
  dplyr::inner_join(pert_bcs) %>%
  reshape2::acast(ccle_name + culture ~ pert, value.var = "LFC_cb") %>%
  cor(use = "p")

# UMAP projection of correlation matrix
pert_umap <- uwot::umap(X = as.dist(1 - pert_cors), n_neighbors = 7) %>%
  as_tibble() %>%
  dplyr::mutate(id = colnames(pert_cors)) %>%
  tidyr::separate(col = "id",
                  into = c("pert_iname", "pert_idose", "x_project_id", "pert_plate"),
                  sep = "::")

# color palette
mycolors <- tableau_color_pal()(10)
mycolors <- colorRampPalette(mycolors)(length(pert_umap$x_project_id %>% unique()))
```

### UMAP

UMAP projection of correlation matrix of compounds.

```{r umap, fig.height=8}
p <- pert_umap %>%
  dplyr::mutate(Perturbation = paste(pert_iname, pert_idose, pert_plate)) %>%
  dplyr::rename(Project = x_project_id) %>%
  ggplot(mapping = aes(x = V1, y = V2, color = Project, name = Perturbation)) +
  labs(x = "UMAP1", y = "UMAP2") +
  geom_point(data = . %>% dplyr::filter(str_detect(Project, "VALIDATION")),
             shape = 17) +
  geom_point(data = . %>% dplyr::filter(str_detect(Project, "VALIDATION", negate = T)),
             shape = 16) +
  scale_color_manual(values = mycolors)

plotly::ggplotly(p)
```

### Heatmap

Heatmap of correlation matrix of compounds.

```{r comp heatmap, fig.height=16, fig.width=16}
projects <- tibble(Project = word(colnames(pert_cors), 3, sep = fixed("::")))
colnames(pert_cors) <- paste(word(colnames(pert_cors), 1, sep = fixed("::")),
                             word(colnames(pert_cors), 4, sep = fixed("::")))
rownames(pert_cors) <- paste(word(rownames(pert_cors), 1, sep = fixed("::")),
                             word(rownames(pert_cors), 4, sep = fixed("::")))

heatmaply::heatmaply(pert_cors,
                     show_dendrogram = F,
                     fontsize_row = 6, fontsize_col = 6,
                     colorbar_yanchor = "top", colorbar_len = 0.1,
                     colorbar_ypos = 1, colorbar_xpos = 1,
                     plot_method = "plotly", limits = c(-1, 1),
                     row_side_colors = projects,
                     row_side_palette = colorRampPalette(mycolors),
                     side_color_colorbar_len = 0.6)
```
